---
title: "Homework #2"
author: "Lynn Chen"
date: "10/4/2021"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```


## Problem 1:

Reading and cleaning **Mr. Trash Wheel** sheet.

```{r trashwheel_df}
trashwheel_df = 
  read_xlsx(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read and clean precipitation data for 2018 and 2019.

```{r 2018 and 2019}
precip_2018 = 
  read_xlsx(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2019 = 
  read_xlsx(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2019) %>% 
  relocate(year)
```

Combine precipitation datasets. 

```{r combine precip}
precip_df = 
  bind_rows(precip_2018, precip_2019) %>% 
  mutate(month = as.character(month))
```

The **Mr. Trash Wheel** data contains information on the date, weight of trash, type of trash collected and the number of houses powered. The number of observations recorded in the **Mr.Trashwheel** and the **precipitation** datasets are `r nrow(trashwheel_df)` and `r nrow(precip_df)`, respectively. 
The key variable associated with precipitation datasets is the precipitation in inches, which quantifies the amount of precipitation recorded. 

* The total precipitation in 2018 is `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. 
* The median number of sports balls in a dumpster in 2019 is `r trashwheel_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.


## Problem 2:

First, clean `pols-month.csv` data. 

```{r pols-month}
pols_month_df = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         month = month.name[month]) %>% 
  mutate(president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
  select(-prez_dem, -prez_gop, -day)
```

Second, clean the data in `snp.csv`.

```{r snp}
snp_df = 
  read_csv("fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>% 
  mutate(date = as.Date(date, format = "%m/%d/%Y")) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(
    year = as.integer(year),
    year = ifelse(year <= 15, year + 2000, year + 1900), 
    month = as.integer(month), 
    day = as.integer(day)) %>%
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day)
```

Third, tidy the `unemployment` data and merge with previous datasets. 

```{r unemployment}
unemployment_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
    ) %>% 
  mutate(month = as.factor(month),
         year = as.integer(year),
         month = as.integer(month),
         month = month.name[month])
```

Merge datasets!!

```{r merge}
pols_snp_df = left_join(pols_month_df, snp_df, by = c("year", "month")) 

pols_snp_unemployment_merge = left_join(pols_snp_df, unemployment_df, by = c("year", "month"))
```

The **pols-month** data contains `r nrow(pols_month_df)` observations of `r ncol(pols_month_df)` variables related to the number of national politicians who are democratic or republican over the given years. 

The **snp** data contains `r nrow(snp_df)` observations of `r ncol(snp_df)` variables related to Standard & Poor’s stock market index from year `r min(select(snp_df,year))` to `r max(select(snp_df,year))`. 

The **unemployment** dataset contains `r nrow(unemployment_df)` observations of `r ncol(unemployment_df)` variables outlining the percentage of unemployment in a given month and year.

Lastly, the merged dataset consists of `r names(pols_snp_unemployment_merge)` variables and contains `r nrow(pols_snp_unemployment_merge)` rows and `r ncol(pols_snp_unemployment_merge)` columns. The data presented is from year `r min(select(pols_snp_unemployment_merge,year))` through `r max(select(pols_snp_unemployment_merge,year))`.  


## Problem 3: 

Load and tidy `popular_baby_names` data.

```{r, message = FALSE}
baby_names_df = 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    ethnicity = recode(
      ethnicity,
      "BLACK NON HISP" = "BLACK NON HISPANIC",
      "WHITE NON HISP" = "WHITE NON HISPANIC", 
      "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER"),
    childs_first_name = str_to_title(childs_first_name),
    gender = str_to_title(gender),
    ethnicity = str_to_title(ethnicity)) %>% 
  distinct()
```

Create a table showing the rank in popularity of the name “Olivia” as a female baby name over time.

```{r olivia table}
olivia_name = 
  filter(
    baby_names_df, 
    childs_first_name == "Olivia",
    gender == "Female") %>% 
  select(-count,-gender,-childs_first_name) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") 

knitr::kable(olivia_name)
```

Create a table showing the most popular name among male children over time.

```{r popular male name table}
popular_male_name = 
  filter(baby_names_df,
         gender == "Male",
         rank == 1) %>% 
  select(-rank,-count,-gender) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name")

knitr::kable(popular_male_name)
```

Create a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) for male, white non-hispanic children born in 2016.

```{r scatter plot}
plot_df = 
  filter(baby_names_df,
         gender == "Male",
         ethnicity == "White Non Hispanic",
         year_of_birth == 2016)
ggplot(plot_df, aes(x = rank, y = count)) + 
geom_point() +
ggtitle("Name ranks of Male White Non-Hispanic children born in 2016") +
xlab("Rank") +
ylab("Count")
```

